EDIpack: A generic and interoperable Lanczos-based solver for quantum impurity problems

### 2025-05-24 04_cbinding.tex
Merge updated from GitHub into OverLeaf. Co-authorhed-by: Adriano Amaricci <adriano.amaricci@gmail.com>

\documentclass[edipack2.tex]{subfiles}
\begin{document}

\section{Interoperability}\label{SecInterop}
The recent growing availability of state-of-the-art software dedicated
to the solution of quantum impurity problems using different methods
poses a serious challenge to test accuracy and realiability of the
results.
As such, complex softwares are requested to develop a higher level of
interoperability, i.e. the possibility to operate with other software
possibly using different programming language.
Modern Fortran, which is the language of choice for \NAME, since many
years supports the standardized generation of procedures and
variables that are interoperable with C.

@@

\subsubsection{Installation and inclusion}\label{sSecInteropCbindingsInstallation}
The C-binding module in included in the build process of \NAME and
compiled into a dynamical library {\tt
  libedipack\_cbindings.so/.dylib}. As dicussed in the
\secu{sSecInstallBuildInstall}, the support for inequivalent
impurities is configured at the build level and propagates to the
C-bindings library as well. An exported variable {\tt has\_ineq} is
defined and export to C/C++ as a way to query the presence of
inequivalent impurities support. 
The generated library and header files get installed in the include
directory at the prefix location, specified during configuration
step. The corresponding path is added to the environment variable {\tt
  LD\_LIBRARY\_PATH}, valid of any Unix/Linux system, via any of the
loading methods outlines in \secu{sSecInstallOSloading}.

The C/C++ compatible functions and variables 
are declared in the header file  {\tt edipack\_cbindings.h}.
Since {\tt ISO\_C\_BINDING} only provides, to date, C compatibility, the 
functions and variables are declared with C linkage, which prevents name mangling.
Function overloading is not supported, hence all interfaced Fortran functions,
for example supporting multiple input variables combinations with different types
and ranks, are here represented by multiple alternative functions.
As an example, the {\tt ed\_chi2\_fitgf} Fortran function, which handles the fitting of bath parameters, is callable in C++ in the following variants
\begin{itemize}
    \item  {\tt chi2\_fitgf\_single\_normal\_n3}: for {\tt ED\_MODE=NORMAL/NONSU2}, single-site DMFT and rank-3 Weiss field/hybridization function arrays
    \item  {\tt chi2\_fitgf\_single\_normal\_n5}: for {\tt ED\_MODE=NORMAL/NONSU2}, single-site DMFT and rank-5 Weiss field/hybridization function arrays
    \item  {\tt chi2\_fitgf\_single\_superc\_n3}: for {\tt ED\_MODE=SUPERC}, single-site DMFT and rank-3 Weiss field/hybridization function arrays
    \item  {\tt chi2\_fitgf\_single\_superc\_n5}: for {\tt ED\_MODE=SUPERC}, single-site DMFT and rank-5 Weiss field/hybridization function arrays
\end{itemize}
and analogous functions for the real-space DMFT case. All functions are listed and documented in the online manual.
When using \NAME functions in a C/C++ program, care must be taken in the way arrays are passed. Consistently with {\tt ISO\_C\_BINDING}, non-scalar parameters have to be passed as raw pointers. An array of integers containing the dimensions of the former need to be passed as well to allow for proper Fortran input parsing. A working C++ example is provided in the {\tt examples} folder of the \NAME repository and discussed in Section \ref{SecExamples}.

\subsubsection{Implementation}\label{sSecInteropCbindingsImplementation}
The interface layer is contained in the Fortran module {\tt
  EDIPACK\_C}. This contains a common part and two set of functions,
one to interface the procedures from \NAME and a second
one to extend the interface to the inequivalent impurities case.
Specifically, the implemented interface functions expose to C through
{\tt bind(C)} statement a number of
procedures composing the Fortran API of \NAME, i.e. contained in {\tt
  ED\_MAIN}. The procedures and shared variables can be divided in four main groups:

\paragraph{{\bf Variables}}
A number of relevant input and shared variables which are normally
required to setup or to control the calculation are interfaced to C
directly in the \NAME modules {\tt ED\_INPUT\_VARIABLES} using {\tt
  bind(C)} constructs. These are implicitly loaded into the C-binding module {\tt
  EDIPACK\_C} through the Fortran {\tt USE EDIPACK} statement and
then further interfaced in the C++ header file.

\paragraph{{\bf Main}} This group contains interface to the exact
digonalization solver, interfacing the solver itself {\tt
  solve\_site/ineq} as well as its initialization
{\tt init\_solver\_site/ineq} and finalization {\tt
  finalize\_solver} procedures. It also includes the functions used to
set the non-interacting part of the impurity Hamiltonian
$h^0_{\a\b\s\s'}$ through the functions {\tt set\_Hloc}, as well as the
interaction Hamiltonian through {\tt add\_twobody\_operator}.

\paragraph{{\bf Bath}} In this group we implement a number of
procedures dealing with bath initialization, symmetry operations and
optimization. In particular it contains the function returning the
dimension of the bath array on the user side {\tt
  get\_bath\_dimension} as well as the setup of the matrix basis
$\vec{\Gamma}$ for the replica/general bath via different instances of
{\tt set\_Hreplica/general\_{site,lattice}\_d{3,5}}. 
The crucial part in the DMFT-Lanczos self-consistency loop is the optimization of the bath
through conjugate-gradient algorithm. To this end we interfaced a
number of functions {\tt
  chi2\_fitgf\_{single,lattice}\_{normal,superc}\_n{3,4,5,6}} which
cover all the cases supported in \NAME. Note that, because the actual
optimization is still performed through the Fortran code, no changes
apply to the outcome of this step.

\paragraph{{\bf Input/Output}}
The input and output part of the software is interfaced in this group
of functions. In particular, {\tt read\_input} expose to C the input
reading procedure of \NAME which set essentially set all the internal
variables of \NAME.
Next, in {\tt edipack{2ineq}\_c\_binding\_io} we interface all the
functions implementing the communication from the \NAME instance to
the user, namely those to retrieve static observables (e.g. {\tt
  get\_dens}), the impurity Green's functions and self-energies
(e.g. {\tt get\_Sigma\_{site,lattice}\_n{3,5}}), the
impurity susceptibilities (e.g. {\tt get\_spinChi}), the impurity reduced density matrix as well
as the non-interacting Green's and hybridization functions starting from the
user bath array.

@@

\subsection{EDIpack2py, the Python API}\label{sSecInteropEDIpy}
As a first application of the \NAME C-bindings we implemented a
complete Python interface called EDIpack2py. This is a Python module which
enables access to all the library features and unlock implementation of
further interfaces of \NAME as a plug-in solver for external Python
based software. Detailed documentation can be found online at \href{https://edipack.github.io/EDIpack2py}
{https://edipack.github.io/EDIpack2py}.

\subsubsection{Installation}\label{sSecInteropEDIpyInstallation}
\paragraph{From source}
EDIpack2py is available as a platform-agnostic Python module depending on EDIpack.
The source code can be obtained from the git repository \href{https://github.com/EDIpack/EDIpack2py}{EDIpack2py} and installed via

@@

\paragraph{PyPi}
The EDIpack2py package is also available in PyPi at
\href{https://pypi.org/project/EDIpack2py/}{pypi.org/EDIpack2py}. The package can be easily installed in any system supporting {\tt pip} as:

@@

\paragraph{Anaconda}
As for \NAME, the Python API in EDIpack2py is available as an
Anaconda package for GNU/Linux and Mac Os systems. Packages are available for
Python version $>3.10$. The \NAME package contains the {\tt
  EDIpack2py} Python module as well as \NAME and SciFortran
libraries. 
%In this case the resolution of the dependencies is
%taken care from Conda itself. 
Using conda or mamba the installation proceeds as following:
\begin{lstlisting}[style=mybash]
conda create -n edipack  #creation of a virtual environment (name can be changed)
conda activate edipack   #activation of the virtual environment
conda install -c conda-forge -c edipack edipack #installation from the edipack channel
\end{lstlisting}

When the{\tt EDIpack2py} module is imported, it attempts to load the dynamic library {\tt
  libedipack\_cbindings.so/.dylib} containing the Fortran-C bindings
for \NAME. By default the library search proceeds as follow: 
\begin{enumerate}
\item First the user can override the location of the library
  (determined during \NAME build configuration) by exporting an
  environment variable called {\tt EDIPACK\_PATH} .
\item By default, the Python module detects the position of the
  Fortran libraries via {\tt pkg-config}. Any of the OS loading method
  outlined in \secu{sSecInstallOSloading} automatically push the
  correct configuration to the {\tt PKG\_CONFIG\_PATH}. 
\item As a last resort the environment variables {\tt
    LD\_LIBRARY\_PATH} and {\tt DYLD\_LIBRARY\_PATH} are analyzed to
  retrieve the correct position. 
\end{enumerate}
If none of the previous attempts succeeds, the module will not not load correctly and an error message will be printed.

@@

\paragraph{{\bf Global variables}}
This includes a subset of the input variables available in \NAME which
are used to control the calculation.
The variables are loaded globally in {\tt EDIpack2py} and can be accessed
or set locally as properties of the class {\tt
  global\_env}. The global variables are initialized, alongside the remaining
default input variables, through a call to procedure {\tt
  edipack2py.global\_env.read\_input} which interface the {\tt
  ed\_read\_input} function in \NAME. 
A given example is reported in the following code extract:

\begin{lstlisting}[language=python,  frame=lines]
import numpy as np
from edipack2py import global_env as ed
ed.Nspin = 1            # set a global variable
mylocalvar = ed.Nspin   # assign to a local variable
print(ed.Nspin)         # all functions can have global variables as arguments
np.arange(ed.Nspin)     # array of integers from 0 to Nspin-1
\end{lstlisting}

\paragraph{{\bf Solver functions}}
This group includes a number of functions enabling initialization,
execution and finalization  of the \NAME solver.
\begin{itemize}
  \item {\tt init\_solver} and {\tt set\_Hloc}. The first 
    initializes the \NAME environment for the quantum impurity problem
    solution, sets the effective bath either reading it from a file or
    initializing it from flat band model. Once this function is
    called, it is not possible to allocate a second instance of the solver.
    {\tt set\_Hloc} sets the
    non-interacting part of the impurity Hamiltonian $\h^0_{\a\b\s\s'}$. 
    Either function take different argument combinations there
    including support for inequivalent impurities.

@@

\paragraph{{\bf Bath functions}}
This set covers the implementation of utility functions handling the
effective bath on the user side as well as interfaces to specific
\NAME procedures, either setting bath properties or applying
conventional symmetry transformation. Here we discuss 
a pair of crucial functions in this group.
\begin{itemize}
\item {\tt bath\_inspect}This function translates between the
  user-accessible continuous bath array and the bath components
  (energy level, hybridization and so on). It functions in both ways,
  given the array returns the components and vice-versa. It
  autonomously determines the type of bath and ED mode.

@@

\paragraph{{\bf Input/Output functions}}
This group includes functions which return to the userspace
observables or dynamical correlation functions evaluated in \NAME and
conserved in the corresponding memory instance. Each function
provides a general interface which encompass all dimension of the
input array there including inequivalent impurities support.
For instance, the function {\tt get\_sigma} returns the self-energy
function array (evaluated on-the-fly) for a specified supported shape,
normal or anomlous type and on a specific axis or frequency domain.

\paragraph{{\bf Auxiliary functions}}
This group include some auxiliary functions, either interfacing \NAME
procedures or defined locally in Python to provide specific new
functionalities. Among the latter we include {\tt get\_ed\_mode} which
return the value of the variable {\tt ed\_mode=normal,superc,nosu2}
and {\tt get\_bath\_type} which similarly returns the value of {\tt bath\_type}.

@@

\subsection{EDIpack2Triqs, the TRIQS interface}\label{sSecInteropTRIQS}
A thin compatibility layer between \NAME and the Toolbox for Research on
Interacting Quantum Systems (TRIQS)\cite{Parcollet2015CPC} called edipack2triqs
is available as a stand-alone project. It is a pure Python package
built upon \NAME's Python API  (Sec.~\ref{sSecInteropEDIpy}) that provides a
limited object-oriented interface to the most important features of \NAME.
edipack2triqs strives to offer seamless interoperability with program tools
based on TRIQS by adopting data types, conventions and usage patterns
common to other TRIQS-based impurity solvers, such as
TRIQS/CTHYB\cite{Seth2016CPC}. It also enables execution of \NAME calculations
in the MPI parallel mode with no extra effort from the user.
Detailed documentation of the package can be 
found online at \href{https://krivenko.github.io/edipack2triqs/}
{https://krivenko.github.io/edipack2triqs/}.

\subsubsection{Installation}\label{sSecInteropTRIQSInstallation}
The package depends on \NAME (which ultimately depends on SciFortran), EDIpack2py
and TRIQS version 3.1 or newer. Assuming the three prerequisites are correctly 
installed in the OS, the current development version of edipack2triqs can be 
installed with {\tt pip} from its 
\href{https://github.com/krivenko/edipack2triqs}{GitHub repository} as follows:

@@

\paragraph{Anaconda}
Another option for installing the package is by using the Anaconda package 
manager on Unix/Linux systems.
The following commands will create a new {\tt conda} environment named 
`edipack' and install the most recently released version of edipack2triqs along 
with its dependencies (\NAME, EDIpack2py and the TRIQS libraries).

@@

\subsubsection{Implementation}\label{sSecInteropTRIQSImplementation}
The programming interface of edipack2triqs is built around the singleton 
Python class {\tt EDIpackSolver} (defined in the module
{\tt edipack2triqs.solver}), whose instance represents the internal state of 
the \NAME library and exposes its functionality through a number of 
attributes and methods.

\paragraph{{\bf Constructor}}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode=normal,superc,nonsu2} (Sec.~\ref{sSecQNs}) and
{\tt bath\_type=normal,hybrid,general} (Sec.~\ref{sSecBath}). In addition, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

\paragraph{{\bf Method {\tt solve}}}
The method {\tt solve()} of {\tt EDIpackSolver} calls
{\tt edipack2py.global\_env.solve()} and therefore performs the bulk of
calculations. It accepts the inverse temperature $\beta$ required to calculate
the expectation value of physical observables, and parameters of energy grids
for Green's function calculations.

\paragraph{{\bf Input parameters}}
It is possible to read off and change parameters of the Hamiltonian between
successive calls to {\tt solve()} via respective attributes of {\tt 
EDIpackSolver}. Note, however, that the changes that would necessitate a change of {\tt ed\_mode} or {\tt bath\_type} are disallowed. The relevant attributes
are the following.
\begin{itemize}
    \item {\tt nspin} --- the number of non-equivalent spin projections 
          (read-only).
    \item {\tt norb} -- the number of impurity orbitals (read-only).
    \item {\tt hloc} --- matrix $h^0_{\alpha\beta\s\s'}$ of
          the quadratic impurity Hamiltonian (\ref{Himp}).
    \item {\tt Uloc}, {\tt Ust}, {\tt Jh}, {\tt Jx}, {\tt Jp} ---
          parameters $U, U', J, J_X \text{ and } J_P$ of the impurity
          interaction term $\hat{H}^\mathrm{int}_\mathrm{imp}$ (\ref{Hint}).
    \item {\tt bath} --- an object representing the bath. Depending on the
          bath type selected upon solver object construction, this object is an
          instance of {\tt BathNormal}, {\tt BathHybrid} or {\tt BathGeneral}
          (all three classes are members of the module
          {\tt EDIpackSolver.bath}). The bath objects support basic arithmetic
          operations so that a mixing scheme within a DMFT calculation can be
          easily implemented. The way to access individual bath parameters is
          specific to each of the three classes, and is described in detail in the       \href{https://krivenko.github.io/edipack2triqs/documentation.html\#module-edipack2triqs.bath}{online API reference}.
\end{itemize}

\paragraph{{\bf Calculation results}}
After a successful invocation of {\tt solve()}, one can extract results of the 
calculation from the following attributes.
\begin{itemize}
    \item {\tt e\_pot}, {\tt e\_kin} --- thermal average of the potential 
          (interaction) and kinetic energy respectively.
    \item {\tt densities}, {\tt double\_occ} --- lists of average densities and
          double occupancies of impurity orbitals.
    \item {\tt magnetization} --- Cartesian components of the average impurity  
          magnetization vectors, one row per orbital.
    \item {\tt superconductive\_phi} --- Matrix of impurity superconductive
          order parameters
          $\phi_{\alpha\beta} = \langle c_{\alpha\up} c_{\beta\down}\rangle$
          in orbital space. {\color{red} Is it defined elsewhere in the text?}
    \item {\tt g\_iw}, {\tt g\_an\_iw} --- Normal and anomalous components of
          the Matsubara impurity Green's function.
    \item {\tt Sigma\_iw}, {\tt Sigma\_an\_iw} --- Normal and anomalous
          components of the Matsubara impurity self-energy.
    \item {\tt g\_w}, {\tt g\_an\_w} --- Normal and anomalous components of
          the real-frequency impurity Green's function.
    \item {\tt Sigma\_w}, {\tt Sigma\_an\_w} --- Normal and anomalous
          components of the real-frequency impurity self-energy.
\end{itemize}

>>>>

\documentclass[edipack_sp.tex]{subfiles}
\begin{document}

\section{Interoperability}\label{SecInterop}
The recent growing availability of state-of-the-art software dedicated
to the solution of quantum impurity problems using different methods \cite{Bulla2008RMP,Parcollet2015CPC,Seth2016CPC,Bauernfeind2017PRX,Ganahl2015PRB,Mejuto_Bath_ASCI-DMFT,...}
poses a serious challenge to test accuracy and reliability of the
results.
As such, complex softwares are requested to develop a higher level of
interoperability, i.e. the possibility to operate with other software
possibly using different programming language.
Modern Fortran, which is the language of choice for \NAME, since many
years supports the standardized generation of procedures and
variables that are interoperable with C.

@@

\subsubsection{Installation and inclusion}\label{sSecInteropCbindingsInstallation}
The C-binding module in included in the build process of \NAME and
compiled into a dynamical library {\tt
  libedipack\_cbindings.so (.dylib)}. As discussed in the
\secu{sSecInstallBuildInstall}, the support for inequivalent
impurities is configured at the build level and propagates to the
C-bindings library as well. An exported variable {\tt has\_ineq} is
defined and export to C/C++ as a way to query the presence of
inequivalent impurities support. 
The generated library and header files get installed in the include
directory at the prefix location, specified during configuration
step. The corresponding path is added to the environment variable {\tt
  LD\_LIBRARY\_PATH}, valid of any Unix/Linux system, via any of the
loading methods outlines in \secu{sSecInstallOSloading}.

The C/C++ compatible functions and variables 
are declared in the header file {\tt e\-di\-pack\_c\-bin\-din\-gs.h}.
Since {\tt ISO\_C\_BINDING} only provides, to date, C compatibility, the 
functions and variables are declared with C linkage, which prevents name mangling.
Function overloading is also not supported, hence all interfaced Fortran functions, for example supporting multiple input variables combinations with different types and ranks, are here represented by multiple alternative functions.
As an example, the {\tt ed\_chi2\_fitgf} Fortran function, which handles the fitting of bath parameters, is callable in C++ in the following variants:
\begin{itemize}
    \item  {\tt chi2\_fitgf\_single\_normal\_n3}: for {\tt ed\_mode=normal/nonsu2}, single-site DMFT and rank-3 Weiss field/hybridization function arrays
    \item  {\tt chi2\_fitgf\_single\_normal\_n5}: for {\tt ed\_mode=normal/nonsu2}, single-site DMFT and rank-5 Weiss field/hybridization function arrays
    \item  {\tt chi2\_fitgf\_single\_superc\_n3}: for {\tt ed\_mode=superc}, single-site DMFT and rank-3 Weiss field/hybridization function arrays
    \item  {\tt chi2\_fitgf\_single\_superc\_n5}: for {\tt ed\_mode=superc}, single-site DMFT and rank-5 Weiss field/hybridization function arrays
\end{itemize}
and analogous functions for the inequivalent impurities case, i.e. EDIpack2Ineq. All functions are thorougly listed and documented in the online manual.
When using \NAME functions in a C/C++ program, care must be taken in the way arrays are passed. Consistently with {\tt ISO\_C\_BINDING}, non-scalar parameters have to be passed as raw pointers. An array of integers containing the dimensions of the former need to be passed as well to allow for proper Fortran input parsing. A working C++ example is provided in the {\tt examples} folder of the \NAME repository and discussed in Section \ref{SecExamples}.

\subsubsection{Implementation}\label{sSecInteropCbindingsImplementation}
The interface layer is contained in the Fortran module {\tt
  EDIPACK\_C}. This contains a common part and two set of functions,
one to interface the procedures from \NAME and a second
one to extend the interface to the inequivalent impurities case.
Specifically, the implemented interface functions expose to C through
{\tt bind(C)} statement a number of
procedures composing the Fortran API of \NAME, i.e. contained in {\tt
  ED\_MAIN}. The procedures and shared variables can be divided in four main groups (note that contrary to the Fortran API the functions here lose the prefix {\tt ed\_}):

\paragraph{{\bf Variables}.}
A number of relevant input and shared variables which are normally
required to setup or to control the calculation are interfaced to C
directly in the \NAME modules {\tt ED\_INPUT\_VARIABLES} using {\tt
  bind(C)} constructs. These are implicitly loaded into the C-binding module {\tt
  EDIPACK\_C} through the Fortran {\tt USE EDIPACK} statement and
then further interfaced in the C++ header file.

\paragraph{{\bf Main}.} This group contains interface to the exact
digonalization solver, interfacing the solver itself {\tt
  solve\_site/ineq} as well as its initialization
{\tt init\_solver\_site/ineq} and finalization {\tt
  finalize\_solver} procedures. It also includes the functions used to
set the non-interacting part of the impurity Hamiltonian
$h^0_{\a\b\s\s'}$ through the functions {\tt set\_Hloc}, as well as the
interaction Hamiltonian through {\tt add\_twobody\_operator}.

\paragraph{{\bf Bath}.} In this group we implement a number of
procedures dealing with bath initialization, symmetry operations and
optimization. In particular it contains the function returning the
dimension of the bath array on the user side {\tt
  get\_bath\_dimension} as well as the setup of the matrix basis
$\vec{\Gamma}$ for the replica (general) bath via different instances of
{\tt set\_Hreplica(general)\_\-\{site,lattice\}\_d\{3,5\}}. 
A crucial part of the DMFT self-consistency loop is the optimization of the bath through CG algorithm (see \secu{sSecFit}). To this end we interfaced a
number of functions {\tt
  chi2\_fitgf\_\{single,lattice\}\_\{normal,superc\}\_n\{3,4,5,6\}} which
cover all the cases supported in \NAME. Note that, because the actual
optimization is still performed through the Fortran code, no changes
apply to the outcome of this step.

\paragraph{{\bf Input/Output}.}
The input and output part of the software is interfaced in this group
of functions. In particular, {\tt read\_input} exposes to C the input
reading procedure of \NAME which set all the internal
variables of \NAME.
Next, in {\tt edipack\{2ineq\}\_c\_binding\_io} we interface all the
functions implementing the communication from the \NAME instance to
the user, namely those to retrieve static observables (e.g. {\tt
  get\_dens}), the impurity Green's functions and self-energies
(e.g. {\tt get\_Sigma\_\{site,lattice\}\_n\{3,5\}}), the
impurity susceptibilities (e.g. {\tt get\_spinChi}), the impurity reduced density matrix as well
as the non-interacting Green's and hybridization functions starting from the user bath array.

@@

\subsection{EDIpack2py, the Python API}\label{sSecInteropEDIpy}
As a first application of the \NAME C-bindings we implemented a
complete Python interface called EDIpack2py. This is a Python module which
enables access to all the library features and unlock implementation of
further interfaces of \NAME as a plug-in solver for external Python
based software. Detailed documentation can be found online at \href{https://edipack.github.io/EDIpack2py}
{edipack.github.io/EDIpack2py}.

\subsubsection{Installation}\label{sSecInteropEDIpyInstallation}
\paragraph{From source.}
EDIpack2py is available as a platform-agnostic Python module depending on EDIpack.
The source code can be obtained from the git repository \href{https://github.com/EDIpack/EDIpack2py}{EDIpack2py} and installed via

@@

\paragraph{PyPi.}
The EDIpack2py package is also available in PyPi at
\href{https://pypi.org/project/EDIpack2py/}{pypi.org/EDIpack2py}. The package can be easily installed in any system supporting {\tt pip} as:

@@

\paragraph{Anaconda.}
As for \NAME, the Python API in EDIpack2py is available as an
Anaconda package for GNU/Linux and Mac Os systems. Packages are available for
Python version $>3.10$. The \NAME package contains the {\tt
  EDIpack2py} Python module as well as \NAME and SciFortran
libraries. 
%In this case the resolution of the dependencies is
%taken care from Conda itself. 
Using Conda or Mamba the installation proceeds as following:
\begin{lstlisting}[style=mybash]
conda create -n edipack  #create a virtual environment
conda activate edipack   #activate it
conda install -c conda-forge -c edipack edipack #install edipack
\end{lstlisting}

\noindent
When the {\tt EDIpack2py}    module is imported, it attempts to load the dynamic library {\tt
  libedipack\-\_cbindings.so (.dylib)} containing the Fortran-C bindings
for \NAME. By default the library search proceeds as follow: 
\begin{enumerate}
\item First the user can override the location of the library
  (determined during \NAME build configuration) by exporting an
  environment variable called {\tt EDIPACK\_PATH} .
\item By default, the Python module detects the position of the
  Fortran libraries via {\tt pkg-config}. Any of the OS loading method
  outlined in \secu{sSecInstallOSloading} automatically push the
  correct configuration to the {\tt PKG\_CONFIG\_PATH}. 
\item As a last resort the environment variables {\tt
    LD\_LIBRARY\_PATH} and {\tt DYLD\_LIBRARY\_PATH} are analyzed to
  retrieve the correct position. 
\end{enumerate}
If none of the previous attempts succeeds, the module will not not load correctly and an error message will be printed.

@@

\paragraph{{\bf Global variables}.}
This includes a subset of the input variables available in \NAME which
are used to control the calculation.
The variables are loaded globally in {\tt EDIpack2py} and can be accessed
or set locally as properties of the class {\tt
  global\_env}. The global variables are initialized, alongside the remaining
default input variables, through a call to procedure {\tt
  edipack2py.global\_env.read\_input} which interface the {\tt
  ed\_read\_input} function in \NAME. 
A given example is reported in the following code extract:

\begin{lstlisting}[style=mypython,frame=lines]
import numpy as np
from edipack2py import global_env as ed
ed.Nspin = 1            # set a global variable
mylocalvar = ed.Nspin   # assign to a local variable
print(ed.Nspin)         # all functions can have global variables as arguments
np.arange(ed.Nspin)     # array of integers from 0 to Nspin-1
\end{lstlisting}

\paragraph{{\bf Solver functions}.}
This group includes a number of functions enabling initialization,
execution and finalization  of the \NAME solver.
\begin{itemize}
  \item {\tt init\_solver} and {\tt set\_Hloc}. The first 
    initializes the \NAME environment for the quantum impurity problem
    solution, sets the effective bath either reading it from a file or
    initializing it from flat band model. Once this function is
    called, it is not possible to allocate a second instance of the solver.
    {\tt set\_Hloc} sets the
    non-interacting part of the impurity Hamiltonian $\h^0_{\a\b\s\s'}$. 
    Either function take different argument combinations there
    including support for inequivalent impurities.

@@

\paragraph{{\bf Bath functions}.}
This set covers the implementation of utility functions handling the
effective bath on the user side as well as interfaces to specific
\NAME procedures, either setting bath properties or applying
conventional symmetry transformation. Here we discuss 
a pair of crucial functions in this group.
\begin{itemize}
\item {\tt bath\_inspect}This function translates between the
  user-accessible continuous bath array and the bath components
  (energy level, hybridization and so on). It functions in both ways,
  given the array returns the components and vice-versa. It
  autonomously determines the type of bath and ED mode.

@@

\paragraph{{\bf Input/Output functions}.}
This group includes functions which return to the userspace
observables or dynamical correlation functions evaluated in \NAME and
conserved in the corresponding memory instance. Each function
provides a general interface which encompass all dimension of the
input array there including inequivalent impurities support.
For instance, the function {\tt get\_sigma} returns the self-energy
function array (evaluated on-the-fly) for a specified supported shape,
normal or anomlous type and on a specific axis or frequency domain.

\paragraph{{\bf Auxiliary functions}.}
This group include some auxiliary functions, either interfacing \NAME
procedures or defined locally in Python to provide specific new
functionalities. Among the latter we include {\tt get\_ed\_mode} which
return the value of the variable {\tt ed\_mode=normal,superc,nosu2}
and {\tt get\_bath\_type} which similarly returns the value of {\tt bath\_type}.

@@

\subsection{EDIpack2Triqs: the TRIQS interface}\label{sSecInteropTRIQS}
A thin compatibility layer between \NAME and the Toolbox for Research on
Interacting Quantum Systems (TRIQS)\cite{Parcollet2015CPC} called EDIpack2Triqs
is available as a stand-alone project. This is a pure Python package
built upon \NAME's Python API EDIpack2py (see \secu{sSecInteropEDIpy}) that provides a
limited object-oriented interface to the most important features of \NAME.
EDIpack2Triqs strives to offer seamless interoperability with program tools
based on TRIQS by adopting data types, conventions and usage patterns
common to other TRIQS-based impurity solvers, such as
TRIQS/CTHYB\cite{Seth2016CPC}. It also enables execution of \NAME calculations
in the MPI parallel mode with no extra effort from the user.
Detailed documentation of the package can be 
found online at \href{https://krivenko.github.io/edipack2triqs/}
{krivenko.github.io/edipack2triqs/}.

\subsubsection{Installation}\label{sSecInteropTRIQSInstallation}
The package depends on \NAME (which ultimately depends on SciFortran), EDIpack2py
and TRIQS version 3.1 or newer. Assuming the three prerequisites are correctly 
installed in the OS, the current development version of EDIpack2triqs can be 
installed with {\tt pip} from its 
\href{https://github.com/krivenko/edipack2triqs}{GitHub repository} as follows:

@@

\paragraph{Anaconda.}
Another option for installing the package is by using the Anaconda package 
manager on Unix/Linux systems.
The following commands will create a new {\tt conda} environment named 
`edipack' and install the most recently released version of EDIpack2triqs along 
with its dependencies (\NAME, EDIpack2py and the TRIQS libraries).

@@

\subsubsection{Implementation}\label{sSecInteropTRIQSImplementation}
The programming interface of EDIpack2triqs is built around the singleton 
Python class {\tt EDIpackSolver}, defined in the module
{\tt edipack2triqs.solver}, whose instance represents the internal state of 
the \NAME library and exposes its functionality through a number of 
attributes and methods.

\paragraph{{\bf Constructor}.}
Constructor of {\tt EDIpackSolver} accepts the Hamiltonian to be diagonalized
in the form of a second-quantized fermionic operator, the TRIQS {\tt Operator}
object described in Sec.~8.7 of Ref.~\onlinecite{Parcollet2015CPC} (impurity 
problems involving bosons are not supported in the current version). In 
addition to the Hamiltonian, four {\it fundamental operator sets}
{\tt fops\_imp\_up}, {\tt fops\_imp\_dn}, {\tt fops\_bath\_up} and
{\tt fops\_bath\_dn} must be provided.
Each of the sets contains pairs of labels {\tt (b,j)} carried
by the operators $c^\dagger_{b,i} / c_{b,i}$ corresponding to either impurity
or bath electronic degrees of freedom with a certain spin projection.
Having these two crucial pieces of information, the constructor initializes
the underlying Fortran library and automatically selects the appropriate
{\tt ed\_mode} (see Sec.~\ref{sSecQNs}) and
{\tt bath\_type} (see Sec.~\ref{sSecBath}). In addition, the
constructor accepts a vast array of keyword arguments that allow for fine-tuning
the diagonalization process. Among others, these include the ED algorithm
selection, the quantum numbers to be used, parameters of the Krylov space,
the spectrum cutoff and various tolerance levels.

\paragraph{{\bf Method {\tt solve}}.}
The method {\tt solve()} of {\tt EDIpackSolver} calls
{\tt edipack2py.global\_env.solve()} and therefore performs the bulk of
calculations. It accepts the inverse temperature $\beta$ required to calculate
the expectation value of physical observables, and parameters of energy grids
for Green's function calculations.

\paragraph{{\bf Input parameters}.}
It is possible to read off and change parameters of the Hamiltonian between
successive calls to {\tt solve()} via respective attributes of {\tt 
EDIpackSolver}. Note, however, that the changes that would necessitate a change of {\tt ed\_mode} or {\tt bath\_type} are disallowed. The relevant attributes
are the following.
\begin{itemize}
    \item {\tt nspin} --- the number of non-equivalent spin projections 
          (read-only).
    \item {\tt norb} -- the number of impurity orbitals (read-only).
    \item {\tt hloc} --- matrix $h^0_{\alpha\beta\s\s'}$ of
          the quadratic impurity Hamiltonian \equ{Himp}.
    \item {\tt Uloc}, {\tt Ust}, {\tt Jh}, {\tt Jx}, {\tt Jp} ---
          parameters $U, U', J, J_X \text{ and } J_P$ of the impurity
          interaction term $\hat{H}^\mathrm{int}_\mathrm{imp}$ \equ{Hint}.
    \item {\tt bath} --- an object representing the bath. Depending on the
          bath type selected upon solver object construction, this object is an
          instance of {\tt BathNormal}, {\tt BathHybrid} or {\tt BathGeneral}
          (all three classes are members of the module
          {\tt EDIpackSolver.bath}). The bath objects support basic arithmetic
          operations so that a mixing scheme within a DMFT calculation can be
          easily implemented. The way to access individual bath parameters is
          specific to each of the three classes, and is described in detail in the       \href{https://krivenko.github.io/edipack2triqs/documentation.html\#module-edipack2triqs.bath}{online API reference}.
\end{itemize}

\paragraph{{\bf Calculation results}.}
After a successful invocation of {\tt solve()}, one can extract results of the 
calculation from the following attributes.
\begin{itemize}
    \item {\tt e\_pot}, {\tt e\_kin} --- thermal average of the potential 
          (interaction) and kinetic energy respectively.
    \item {\tt densities}, {\tt double\_occ} --- lists of average densities and
          double occupancies of impurity orbitals.
    \item {\tt magnetization} --- Cartesian components of the average impurity  
          magnetization vectors, one row per orbital.
    \item {\tt superconductive\_phi} --- Matrix of impurity superconductive
          order parameters
          $\phi_{\alpha\beta} = \langle c_{\alpha\up} c_{\beta\down}\rangle$
          in orbital space.
    \item {\tt g\_iw}, {\tt g\_an\_iw} --- Normal and anomalous components of
          the Matsubara impurity Green's function.
    \item {\tt Sigma\_iw}, {\tt Sigma\_an\_iw} --- Normal and anomalous
          components of the Matsubara impurity self-energy.
    \item {\tt g\_w}, {\tt g\_an\_w} --- Normal and anomalous components of
          the real-frequency impurity Green's function.
    \item {\tt Sigma\_w}, {\tt Sigma\_an\_w} --- Normal and anomalous
          components of the real-frequency impurity self-energy.
\end{itemize}
